\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-05-13 paper.qmd
Add best model but getting an error because r2 = 0.59 something working better when r2 > 1

The value for $b_2$ is computed by fixing $b_1$. Considering the lower bound of the NLDR, $a_1 > 2 \times s_1$, and $a_1 > \frac{1-s_1}{b_1 -1}$. Similarly, according to the upper bound of the NLDR, $a_1 > \frac{2(r_2 - s_2)}{\sqrt{3}(b_2 - 1)}$, because $a_2 = \frac{\sqrt(3)}{2}a_1$ for regular hexagons. The $a_1$ that satisfies the constraints will be selected. Therefore, $b_2 = \Big\lceil1 +\frac{2(r_2 - s_2)(b_1 - 1)}{\sqrt{3}(1-s_1)}\Big\rceil$.

@@

pacmap_minst <- read_rds("data/mnist/mnist_pacmap.rds")
pacmap_minst_scaled <- gen_scaled_data(data = pacmap_minst)$scaled_nldr

@@

<!-- Fit the model and compute error-->
```{r}

```

@@

imge_error_sample_within <- pixels_gathered |> 
  ggplot(aes(x, y, fill = value)) +  
  geom_tile() +  
  facet_wrap(~ factor(instance, levels = c( 4699, 4785, 5173, 6671, 2186, 4636, 
                                            4882, 773, 2825, 6169, 6948, 5987, 
                                            1282, 7715, 622, 2989, 6912, 4370, 
                                            1227, 4042, 2444, 836, 1654)), 
             ncol = 12) +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank())
```

<!--outside the nonlinear structure-->
```{r}
#| eval: false

img_error <- error_df |> 
  filter(total >= 8) |> 
  filter((!ID %in% c(4699, 4785, 5173, 6671, 2186, 4636, 4882, 773, 
                            2825, 6169, 6948, 5987, 1282, 7715, 622, 2989, 
                            6912, 4370, 1227, 4042, 2444, 836, 1654))) |>
  pull(ID)

@@

imge_error_sample <- pixels_gathered |> 
  ggplot(aes(x, y, fill = value)) +  
  geom_tile() +  
  facet_wrap(~ instance, ncol = 11) +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank())
```

```{r}
#| eval: false
##| echo: false
#| fig-cap: "(a) Images of handwritten digit 1 which occur large model error within the nonlinear strcuture, and (b) Images of handwritten digit 1 which occur large model error outside the nonlinear structure."
#| label: fig-anomalies
#| out-width: 100%
#| fig-pos: H

imge_error_sample_within + imge_error_sample +
  plot_annotation(tag_levels = 'a')
```

>>>>

The value for $b_2$ is computed by fixing $b_1$. Considering the lower bound of the NLDR, $a_1 > 2 \times s_1$, and $a_1 > \frac{1-s_1}{b_1 -1}$. Similarly, according to the upper bound of the NLDR, $a_1 > \frac{2(r_2 - s_2)}{\sqrt{3}(b_2 - 1)}$, because $a_2 = \frac{\sqrt(3)}{2}a_1$ for regular hexagons. Therefore, $b_2 = \Big\lceil1 +\frac{2(r_2 - s_2)(b_1 - 1)}{\sqrt{3}(1-s_1)}\Big\rceil$.

@@

pacmap_minst <- read_rds("data/mnist/mnist_pacmap.rds") |> 
  select(PaCMAP1, PaCMAP2, ID)
mnist_scaled_obj <- gen_scaled_data(
  data = pacmap_minst)
pacmap_minst_scaled <- mnist_scaled_obj$scaled_nldr

@@

<!-- Fit the model and compute error-->
```{r}
## Compute hexbin parameters
num_bins_x_mnist <- 20
lim1 <- mnist_scaled_obj$lim1
lim2 <- mnist_scaled_obj$lim2
r2_mnist <- diff(lim1)/diff(lim2) # Need to correct r2_mnist <- diff(lim2)/diff(lim1)

mnist_model <- fit_highd_model(
  training_data = training_data_mnist,
  emb_df = pacmap_minst_scaled,
  bin1 = num_bins_x_mnist,
  q = 0.1,
  r2 = r2_mnist,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "PC"
)

df_bin_centroids_mnist <- mnist_model$df_bin_centroids
df_bin_mnist <- mnist_model$df_bin

## Triangulate bin centroids
tr1_object_mnist <- tri_bin_centroids(
  df_bin_centroids_mnist, x = "c_x", y = "c_y")
tr_from_to_df_mnist <- gen_edges(
  tri_object = tr1_object_mnist)

## Compute 2D distances
distance_mnist <- cal_2d_dist(
  tr_coord_df = tr_from_to_df_mnist, 
  start_x = "x_from", 
  start_y = "y_from", 
  end_x = "x_to", 
  end_y = "y_to", 
  select_vars = c("from", "to", "distance"))

## To find the benchmark value
benchmark_mnist <- find_lg_benchmark(
  distance_edges = distance_mnist, 
  distance_col = "distance")

trimesh_removed_mnist <- vis_rmlg_mesh(
  distance_edges = distance_mnist, 
  benchmark_value = benchmark_mnist, 
  tr_coord_df = tr_from_to_df_mnist, 
  distance_col = "distance") +
  #xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("a", sc_ltr_pos)

## Compute error
error_df <- augment(
  df_bin_centroids = df_bin_centroids_mnist, 
  df_bin = df_bin_mnist, 
  training_data = training_data_mnist, 
  newdata = NULL, 
  type_NLDR = "PaCMAP", 
  col_start = "PC")

## Categorize error

error_df <- error_df |>
  mutate(type = case_when(
    row_wise_abs_error <= 2 ~ "error 0-2",
    row_wise_abs_error <= 4 ~ "error 2-4",
    row_wise_abs_error <= 6 ~ "error 4-6",
    row_wise_abs_error <= 6 ~ "error 6-8",
    row_wise_abs_error <= 10 ~ "error 8-10",
    .default = "error greter than 10"
  )) |>
  mutate(type = factor(type, levels = c(
    "error 0-2", "error 2-4", "error 4-6", "error 6-8", "error 8-10", 
    "error greter than 10")))

## To join embedding
error_df <- error_df |>
  bind_cols(pacmap_minst |> 
              select(-ID))

error_plot_mnist <- error_df |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2, 
             color = type, 
             group = ID)) +
  geom_point(alpha=0.5, 
             size = 0.1) +
  interior_annotation("b", sc_ltr_pos)
```

```{r}
#| echo: false
#| fig-cap: "(a) Model generated in the 2D space overlaid on PaCMAP data, and (b) high-D model error in model space."
#| label: fig-model-mnist
#| fig-pos: H

trimesh_removed_mnist + error_plot_mnist + 
  plot_layout(guides='collect', ncol=2) &
  theme(legend.position='bottom')
```

@@

imge_error_sample_within <- pixels_gathered |> 
  ggplot(aes(x, y, fill = value)) +  
  geom_tile() +  
  facet_wrap(~ factor(instance, levels = c( 4699, 4785, 5173, 6671, 2186, 4636, 
                                            4882, 773, 2825, 6169, 6948, 5987, 
                                            1282, 7715, 622, 2989, 6912, 4370, 
                                            1227, 4042, 2444, 836, 1654)), 
             ncol = 5) +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank())
```

<!--outside the nonlinear structure-->
```{r}
img_error <- error_df |> 
  filter(row_wise_abs_error >= 8) |> 
  filter((!ID %in% c(4699, 4785, 5173, 6671, 2186, 4636, 4882, 773, 
                            2825, 6169, 6948, 5987, 1282, 7715, 622, 2989, 
                            6912, 4370, 1227, 4042, 2444, 836, 1654))) |>
  pull(ID)

@@

imge_error_sample <- pixels_gathered |> 
  ggplot(aes(x, y, fill = value)) +  
  geom_tile() +  
  facet_wrap(~ instance, ncol = 5) +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank())
```

```{r}
#| echo: false
#| fig-cap: "(a) Images of handwritten digit 1 which occur large model error within the non-linear strcuture, and (b) Images of handwritten digit 1 which occur large model error outside the non-linear structure."
#| label: fig-anomalies
#| fig-width: 4
#| fig-height: 4
#| out-width: 100%
#| fig-pos: H

imge_error_sample_within + imge_error_sample +
  plot_layout(ncol=2) +
  plot_annotation(tag_levels = 'a')
```

### 2024-05-13 paper.tex
Add best model but getting an error because r2 = 0.59 something working better when r2 > 1

The value for \(b_2\) is computed by fixing \(b_1\). Considering the
lower bound of the NLDR, \(a_1 > 2 \times s_1\), and
\(a_1 > \frac{1-s_1}{b_1 -1}\). Similarly, according to the upper bound
of the NLDR, \(a_1 > \frac{2(r_2 - s_2)}{\sqrt{3}(b_2 - 1)}\), because
\(a_2 = \frac{\sqrt(3)}{2}a_1\) for regular hexagons. The \(a_1\) that
satisfies the constraints will be selected. Therefore,
\(b_2 = \Big\lceil1 +\frac{2(r_2 - s_2)(b_1 - 1)}{\sqrt{3}(1-s_1)}\Big\rceil\).

@@

>>>>

The value for \(b_2\) is computed by fixing \(b_1\). Considering the
lower bound of the NLDR, \(a_1 > 2 \times s_1\), and
\(a_1 > \frac{1-s_1}{b_1 -1}\). Similarly, according to the upper bound
of the NLDR, \(a_1 > \frac{2(r_2 - s_2)}{\sqrt{3}(b_2 - 1)}\), because
\(a_2 = \frac{\sqrt(3)}{2}a_1\) for regular hexagons. Therefore,
\(b_2 = \Big\lceil1 +\frac{2(r_2 - s_2)(b_1 - 1)}{\sqrt{3}(1-s_1)}\Big\rceil\).

@@

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-mnist-1.pdf}

\caption{\label{fig-model-mnist}(a) Model generated in the 2D space
overlaid on PaCMAP data, and (b) high-D model error in model space.}

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-anomalies-1.pdf}

\caption{\label{fig-anomalies}(a) Images of handwritten digit 1 which
occur large model error within the non-linear strcuture, and (b) Images
of handwritten digit 1 which occur large model error outside the
non-linear structure.}
